---
title: "Stats-413-HW6"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r lib}
library(faraway)
library(ggplot2)
library(FNN)
suppressMessages(library(dplyr))
suppressMessages(library(GGally))
suppressMessages(library(MASS))
suppressMessages(library(corrplot))
library(ISLR2)
```

## Q5a.
```{r}
data("prostate")
p_lm = lm(lpsa~. , data=prostate)
summary(p_lm)

x = model.matrix(p_lm)[,-1]
e = eigen(t(x) %*% x)
e$val

sqrt(e$val[1]/e$val)
```
There is a wide range in the eigenvalues and several condition numbers are large. \newline
This means that problems are being caused by more than just one linear combination.

## Q5b.
```{r}
round(cor(prostate[,-9]),2)
```
The variables that are strongly correlated with each other are: \newline
1. svi and lcp, \newline
2. lcp and lcavol, \newline
3. gleason and pgg45

## Q5c.
```{r}
vif(p_lm)
```
There doesn't seem to be any inflation of variance. Thus, I can conclude that there is no evidence that col-linearity is causing predictors to be significant.

## Q6a.
```{r}
data("cheddar")
c_lm = lm(taste~., data=cheddar)
summary(c_lm)
```
Looking at the p-value of Lactic, Lactic is statistically significant in this model at significance level=0.05.

## Q6b.
```{r}
summary = summary(c_lm)
p_value = summary$coef["Lactic","Pr(>|t|)"]
p_value
```

## Q6c.
```{r}
cheddar_new = cheddar
n = nrow(cheddar_new)
cheddar_new$Lactic_new = cheddar_new$Lactic + rnorm(n, mean = 0, sd = 0.01)

##refit the model
fit_new = lm(taste~ Acetic + H2S + Lactic_new, data = cheddar_new)
fit_summary = summary(fit_new)
fit_summary

p = fit_summary$coef["Lactic_new","Pr(>|t|)"]
paste("The p-value of for the previous test is", p)
```

## Q6d.
```{r}
p_list <- rep(0,1000)

for (i in 1:1000){
cheddar_new$Lactic_new <- cheddar_new$Lactic + rnorm(n, mean = 0, sd = 0.01)
fit_new = lm(taste~ Acetic + H2S + Lactic_new, data = cheddar_new)
fit_summary = summary(fit_new)
p_list[i] = fit_summary$coef["Lactic_new","Pr(>|t|)"]
}

mean(p_list)
```
The measurement error did not make enough of a difference to the p-value to make a significant change to the conclusions.

## Q6e.
```{r}
p_list <- rep(0,1000)

for (i in 1:1000){
cheddar_new$Lactic_new <- cheddar_new$Lactic + rnorm(n, mean = 0, sd = 0.1)
fit_new = lm(taste~ Acetic + H2S + Lactic_new, data = cheddar_new)
fit_summary = summary(fit_new)
p_list[i] = fit_summary$coef["Lactic_new","Pr(>|t|)"]
}

mean(p_list)
```
The change of standard deviation from 0.01 to 0.1 has increased the mean p-value. Since the p-value has increased significantly higher than 0.05, the measurement error make a difference.

